Giant Low Surface Brightness Galaxies As Merger Remnants

### 2016-07-29 glsb.tex
edits

By eye, GLSB systems resemble large faint spiral disks at large radii, and regular ellipticals at their centers.  In the Sloan Digital Sky Survey Data Relase 10 (SDSS DR10) \citep{Ahn14}, the Galaxy Zoo classifications \citep{Willett13} are spiral for Malin 1 and 2 and UGC 6614.  In this section, we compare the properties of the GLSB centers and disks with other systems.

@@

The central regions of GLSBs are often refered to in the literature as ``bulges'', however, looking at Figure~\ref{elliptical_relations}, it is clear that these are much brighter than bulges of typical spiral galaxies and the centers of GLSBs are on the bright end of the red-sequence.

@@

The diameters of GLSB disks are very large, with Malin~1's stellar disk detected out to a radius of 80 kpc \citep{Galaz15}.  Stellar disks are typically embedded in much larger \HI\ disks.  In Figure~\ref{disk_comp}, we compare the sizes of the stellar and \HI\ disks of the GLSB galaxies with other observations.  For this plot, we take the maximum detected radius from the literature for the galaxies.  It is possible that deeper observations could increase the values of $R_{max}$.  While the stellar disk of UGC~6614 is close in size to local observed \HI\ disks, Malin~1 and~2 are larger than hte \HI\ disks observed in the THINGS survey.  This seems to point against the GLSB formation mechanism of \citet{Mapelli08}, as we do not observe \HI\ disks in normal spirals large enough to convert into GLSB systems unless an interaction is able to also boost the angular momentum of the resulting stellar disk.

% /Users/yoachim/Research/GLSB/THINGS/thingsplot.py
\begin{figure*}
\epsscale{.5}
\plotone{Plots/things.eps}
\epsscale{1}
\caption{ Comparison of the GLSB disks with other galaxy disks from the literature.  THINGS data taken from \citet{DeBlok2008}, GLSB \HI\ data from Malin 1 from \citet{Pickering97} and GLSB stellar radii from \citet{Galaz15}, \citet{Kasparova14}, and this paper.  Here, $R_{max}$ is the maximum radius with an \HI\ or optical detection.  For the GLSB galaxies, the solid points show the maximum \HI\ detection radius and stars-shaped points show the maximum detected stellar light radius.  All points are maximum detection radii, and thus should be considered lower limits.   \label{disk_comp}}
\end{figure*}

@@

Figures~\ref{malin2_binmap} and~\ref{ugc_binmap} show the fibers that were binned and the resulting spectra for Malin 2 and UGC 06614.  Both galaxies have elliptical-galaxy like spectra in their centers, with outskirts that have significant emission lines.  We find the best-fitting single model spetrum for each observed binned spectrum. When fitting, we include a low order multiplicitave polynomial to match any dust extinction or flux calibration errors in the observations.  This means our fits are for the relative star formation histories, but not the absolute star formation rates (e.g., a model constant star formation history of 1 M$_\sun$ yr$^{-1}$ is an equally good fit as 10 M$_\sun$ yr$^{-1}$.)

@@

An accretion formation scenario does the best job explaining all the observations.  While a collisional ring could fade to a GLSB, the stellar population should be relatively young, and one would not necissarily expect a steep metallicity gradient.  Also, there are no obvious galaxies near GLSBs to act as progenitor colliders.

@@

\citet{Wilkinson15} find -0.15 dex/$R_e$ gradients for early type galaxies and flat metallitcity gradients for late type galaxies.

@@

\citet{Hagen16} present the discovery of a GLSB system around UGC~1382.  As in this work, they find the best explination of the extended disk as being the remnants of an accreted drawf galaxy similar to the simulations of \citet{Pen06}.

@@

We have observed two giant low surface brightness galaxies with the VIRUS-P IFU.  We find the systems are best explained as massive elliptical galaxies which have managed to accrete stars and gas into a large disk.  We find the inner bulge is similar to elliptical galaxies found in SDSS.  The stars have a disk-like velocity field, are dominated by old stars, but do show signs of ongoing star formation. The stars also show a sharp metallicity transition, consistent with the disk forming independently of the bulge. The \HI content of the disk is also consistent with coming from a low-mass accreted galaxy.

We are currently observing higher spatial resolution IFU data with Gemini (GN-2015A-Q-75) of the central region of these galaxies to confirm they are normal elliptical galaxies and to help constrain a more sophisticated dynamical model.

%--------------------------------------
\acknowledgments
Thanks to Dave Doss and the staff at McDonald Observatory for their help on these observing runs. Thanks to Dmitry Bizyaev for stimulating conversations about gLSBs. Thanks to Lauren MacArthur for discussions on stellar metallicity gradients.
%Thanks to the observing specialists at Apache Point for their help executing our remote observations.

@@

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
%\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Upper panels show kinematics from individual fibers, lower panels show the results after fibers have been binned into tripples.  Left:  Stellar velocities measured with pPXF \citep{Cappellari04}. Middle:  Best fitting tilted ring model from kinemetry \citet{Krajn06}. Right: Best fit residuals.  \label{malin_kinem_fibers}}
\end{figure*}

\begin{figure*}
\epsscale{0.25}
%\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
%\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Rotation curves extracted from the best fitting model in Fig~\ref{malin_kinem_fibers}.  The dashed curve shows the expected stellar mass contribution to the rotation curve.  Vertical lines mark $R_e$ and $5R_e$.  \HI\ data from \citet{Pickering97}.  \label{malin_kinem}}
\end{figure*}

@@

\begin{figure}
\plottwo{Plots/Malin2bin_map.eps}{Plots/Malin2binned_spec.eps}
\caption{Left:  Map of which fibers were binned together.  Right:  Resulting binned spectra, residuals from the bright skyline at 5577\AA\ have been masked. \label{malin2_binmap}}
\end{figure}

@@

\begin{figure}
\epsscale{.5}
\plotone{Plots/Malin2_chibar_sfr.eps}
\epsscale{1}
\caption{ Best fitting star formation histories.  For each bin, the SFH is normalized by the peak at each radius.  Error bars are based on $\Delta\chi^2=100$ or the best fitting 4 models, whichever provides more points.  \label{sfh_malin2} }
\end{figure}

\begin{figure}
\epsscale{.5}
\plotone{Plots/Malin2_chibar_z.eps}
\epsscale{1}
\caption{Best fitting stellar metallicities.   Dashed line shows the maximum model metallicity used in the models (minimum is below the y range).  Error bars are as in Figure~\ref{sfh_malin2}. \label{metal_malin2}}
\end{figure}

@@

\begin{figure}
\plottwo{Plots/UGC06614bin_map.eps}{Plots/UGC06614binned_spec.eps}
\caption{Left:  Map showing which fibers were binned for UGC 6614.  Right:  The binned spectra, color-coded to match the map on the left. \label{ugc_binmap}}
\end{figure}

@@

\begin{figure}
\epsscale{.5}
\plotone{Plots/UGC06614_chibar_sfr.eps}
\caption{ Best fitting star formation histories.  For each bin, the SFH is normalized by the peak bin.  Error bars are based on $\Delta\chi^2=100$ or the 4 best fitting models, whichever provides more points.  \label{sfh_ugc} }
\epsscale{1}
\end{figure}

\begin{figure}
\epsscale{.5}
\plotone{Plots/UGC06614_chibar_z.eps}
\epsscale{1}
\caption{Best fitting stellar metallicities.   \label{metal_ugc}}
\end{figure}

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataUGC06614.eps}\plotone{Plots/modelUGC06614.eps}\plotone{Plots/data_modelUGC06614.eps} \\
\plotone{Plots/databinned_UGC06614.eps}\plotone{Plots/modelbinned_UGC06614.eps}\plotone{Plots/data_modelbinned_UGC06614.eps}
%\epsscale{.5}
\epsscale{1}
\plottwo{Plots/kinembinned_UGC06614.eps}{Plots/kinem_HI_binned_UGC06614.eps}
\caption{XXX-break into 2 figures! Fitting the velocity maps with kinemetry package. Vertical lines mark $R_e$ and $5R_e$.  The dashed lines shows a rotation curve based in the enclosed flux, normalized so that all the SDSS stellar mass is contained within $3R_e$.  \HI\ data from \citet{Pickering97}.\label{ugc_kinem}}
\end{figure*}

@@

% LocalWords:  Debattista Horrocks Lancashire RCUK Malin Stanonik Spavone Lelli
% LocalWords:  Bothun GLSB Das Mapelli Weij Loebman SungWon Kwak Scd Sd
% LocalWords:  Schiminovich Iodice Belokurov Fairchild RMS pPXF Cappellari th
% LocalWords:  GANDALF Sarzi CosmoloPy Bernardi GLSBs DeBlok Verheijen Yoachim
% LocalWords:  SEDs Brookhaven Instituto Astrofisica Canarias Notre JINA DM
% LocalWords:  IFU UGC HSB SAURON ellipticals Spitzer IC Galaz
% LocalWords:  kpc HIPASS CCD metallicity MANGA SDSS arcsec lensing IRAF
% LocalWords:  Spectrograph spectrograph IDL longslit photometric FWHM SSP mis
% LocalWords:  dispersions Balmer decrement fiducia redshift Chabrier
% LocalWords:  Faber redshifts NGC kinemetry photometry anisotropy Gyr SFH
% LocalWords:  metallicities colliders rotators et al MNRAS de datacube Ppxf
% LocalWords:  dashed CDM Kasparova LSB Dutton Raskutti Sa Morganti
%%  LocalWords:  Emonts Weijmans Ibata McConnachie Schweizer Magain
%%  LocalWords:  Oosterloo baryonic Buote GALEX Salim Kelson outlier
%%  LocalWords:  Bruzual Ahn Willett SQL Chilingarian Saulder Swaters
%%  LocalWords:  Krajn GN Kuntschner situ indices Emsell Capp CasJobs
%%  LocalWords:  Schmitz circ accrete jabref CMD

>>>>

By eye, GLSB systems resemble large faint spiral disks at large radii, and regular ellipticals at their centers.  In the Sloan Digital Sky Survey Data Release 10 (SDSS DR10) \citep{Ahn14}, the Galaxy Zoo classifications \citep{Willett13} are spiral for Malin 1 and 2 and UGC 6614.  In this section, we compare the properties of the GLSB centers and disks with other systems.

@@

The central regions of GLSBs are often referred to in the literature as ``bulges'', however, looking at Figure~\ref{elliptical_relations}, it is clear that these are much brighter than bulges of typical spiral galaxies and the centers of GLSBs are on the bright end of the red-sequence.

@@

The diameters of GLSB disks are very large, with Malin~1's stellar disk detected out to a radius of 80 kpc \citep{Galaz15}.  Stellar disks are typically embedded in much larger \HI\ disks.  In Figure~\ref{disk_comp}, we compare the sizes of the stellar and \HI\ disks of the GLSB galaxies with other observations.  For this plot, we take the maximum detected radius from the literature for the galaxies.  It is possible that deeper observations could increase the values of $R_{max}$.  While the stellar disk of UGC~6614 is close in size to local observed \HI\ disks, Malin~1 and~2 are larger than the \HI\ disks observed in the THINGS survey.  This seems to point against the GLSB formation mechanism of \citet{Mapelli08}, as we do not observe \HI\ disks in normal spirals large enough to convert into GLSB systems unless an interaction is able to also boost the angular momentum of the resulting stellar disk.

% /Users/yoachim/Research/GLSB/THINGS/thingsplot.py
\begin{figure*}
\epsscale{.5}
\plotone{Plots/things.eps}
\epsscale{1}
\caption{ Comparison of the GLSB disks with other galaxy disks from the literature.  THINGS data taken from \citet{DeBlok2008}, GLSB \HI\ data from Malin 1 from \citet{Pickering97} and GLSB stellar radii from \citet{Galaz15}, \citet{Kasparova14}, and this paper.  Here, $R_{max}$ is the maximum radius with an \HI\ or optical detection.  For the GLSB galaxies, the circular points show the maximum \HI\ detection radius and stars-shaped points show the maximum detected stellar light radius.  All points are maximum detection radii, and thus should be considered lower limits.   \label{disk_comp}}
\end{figure*}

@@

Figures~\ref{malin2_binmap} and~\ref{ugc_binmap} show the fibers that were binned and the resulting spectra for Malin 2 and UGC 06614.  Both galaxies have elliptical-galaxy like spectra in their centers, with outskirts that have significant emission lines.  We find the best-fitting single model spectrum for each observed binned spectrum. When fitting, we include a low order multiplicative polynomial to match any dust extinction or flux calibration errors in the observations.  This means our fits are for the relative star formation histories, but not the absolute star formation rates (e.g., a model constant star formation history of 1 M$_\sun$ yr$^{-1}$ is an equally good fit as 10 M$_\sun$ yr$^{-1}$.)

@@

An accretion formation scenario does the best job explaining all the observations.  While a collisional ring could fade to a GLSB, the stellar population should be relatively young, and one would not necessarily expect a steep metallicity gradient.  Also, there are no obvious galaxies near GLSBs to act as progenitor colliders.

@@

\citet{Wilkinson15} find -0.15 dex/$R_e$ gradients for early type galaxies and flat metallicity gradients for late type galaxies.

@@

\citet{Hagen16} present the discovery of a GLSB system around UGC~1382.  As in this work, they find the best explanation of the extended disk as being the remnants of an accreted dwarf galaxy similar to the simulations of \citet{Pen06}.

@@

We have observed two giant low surface brightness galaxies with the VIRUS-P IFU.  We find the systems are best explained as massive elliptical galaxies which have managed to accrete stars and gas into a large disk.  We find the inner bulge is similar to elliptical galaxies found in SDSS.  The extended galaxy stars have disk-like velocity fields, are dominated by old stars, but do show signs of ongoing star formation. The stars also show a sharp metallicity transition, consistent with the disk forming independently of the bulge. The \HI\ mass of the disk is also consistent with coming from a low-mass accreted galaxy.

We are currently observing higher spatial resolution IFU data with Gemini (GN-2015A-Q-75) of the central region of these galaxies to confirm they are normal elliptical galaxies and constrain a more sophisticated dynamical model of the system.

%--------------------------------------
\acknowledgments
Thanks to Dave Doss and the staff at McDonald Observatory for their help on these observing runs. Thanks to Dmitry Bizyaev for stimulating conversations about GLSBs. Thanks to Lauren MacArthur for discussions on stellar metallicity gradients.
%Thanks to the observing specialists at Apache Point for their help executing our remote observations.

@@

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
%\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Upper panels show kinematics from individual fibers, lower panels show the results after fibers have been binned into triples.  Left:  Stellar velocities measured with pPXF \citep{Cappellari04}. Middle:  Best fitting tilted ring model from kinemetry \citet{Krajn06}. Right: Best fit residuals.  \label{malin_kinem_fibers}}
\end{figure*}

\begin{figure*}
\epsscale{0.25}
%\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
%\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Rotation curves extracted from the best fitting models in Fig~\ref{malin_kinem_fibers}.  The dashed curves show the expected stellar mass contribution to the rotation curve.  Vertical lines mark $R_e$ and $5R_e$.  \HI\ data from \citet{Pickering97}.  \label{malin_kinem}}
\end{figure*}

@@

\begin{figure}
\plottwo{Plots/Malin2bin_map.eps}{Plots/Malin2binned_spec.eps}
\caption{Left:  Map of which fibers were binned together for star formation history measurements.  Right:  Resulting binned spectra, residuals from the bright skyline at 5577\AA\ have been masked.  The inner bins show strong Mg, Fe, and Ca absorption features, while the outer bins have \hb\ and \ion{O}{3} emission lines. \label{malin2_binmap}}
\end{figure}

@@

\begin{figure}
\epsscale{.5}
\plotone{Plots/Malin2_chibar_sfr.eps}
\epsscale{1}
\caption{ Best fitting star formation histories.  For each bin, the SFH is normalized by the peak at each radius.  Error bars are based on $\Delta\chi^2=100$ or the best fitting 4 models, whichever provides more points.  The oldest age bin dominates at all radii.  \label{sfh_malin2} }
\end{figure}

\begin{figure}
\epsscale{.5}
\plotone{Plots/Malin2_chibar_z.eps}
\epsscale{1}
\caption{Best fitting stellar metallicities.   Dashed line shows the maximum model metallicity used in the models (minimum is below the y range limit).  Error bars are as in Figure~\ref{sfh_malin2}.  \label{metal_malin2}}
\end{figure}

@@

\begin{figure}
\plottwo{Plots/UGC06614bin_map.eps}{Plots/UGC06614binned_spec.eps}
\caption{Left:  Map showing which fibers were binned for UGC 6614.  Right:  The binned spectra, color-coded to match the map on the left.  As with Malin~2, the inner spectra show strong metallicity absorption features while the outer spectra show emission lines. \label{ugc_binmap}}
\end{figure}

@@

\begin{figure}
\epsscale{.5}
\plotone{Plots/UGC06614_chibar_sfr.eps}
\caption{Best fitting star formation histories.  For each bin, the SFH is normalized by the peak bin.  Error bars are based on $\Delta\chi^2=100$ or the 4 best fitting models, whichever provides more points.  As with Malin~2, old stars dominate at all radii.  \label{sfh_ugc} }
\epsscale{1}
\end{figure}

\begin{figure}
\epsscale{.5}
\plotone{Plots/UGC06614_chibar_z.eps}
\epsscale{1}
\caption{Best fitting stellar metallicities for the binned UGC~6614 spectra.  Dashed line shows the maximum model value.  \label{metal_ugc}}
\end{figure}

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataUGC06614.eps}\plotone{Plots/modelUGC06614.eps}\plotone{Plots/data_modelUGC06614.eps} \\
\plotone{Plots/databinned_UGC06614.eps}\plotone{Plots/modelbinned_UGC06614.eps}\plotone{Plots/data_modelbinned_UGC06614.eps}
\epsscale{1}
\caption{Upper panels show kinematics from individual fibers, lower panels show the results after fibers have been binned into triples.  Left:  Stellar velocities measured with pPXF \citep{Cappellari04}. Middle:  Best fitting tilted ring model from kinemetry \citet{Krajn06}. Right: Best fit residuals.  \label{ugc_kinem_fibers}}
\end{figure*}

\begin{figure*}
\epsscale{1}
\plottwo{Plots/kinembinned_UGC06614.eps}{Plots/kinem_HI_binned_UGC06614.eps}
\caption{Rotation curves extracted from the best fitting model in Fig~\ref{ugc_kinem_fibers}.  The dashed curve shows the expected stellar mass contribution to the rotation curve.  Vertical lines mark $R_e$ and $5R_e$.  \HI\ data from \citet{Pickering97}. \label{ugc_kinem}}
\end{figure*}

@@

% LocalWords:  Debattista Horrocks Lancashire RCUK Malin Stanonik Spavone Lelli
% LocalWords:  Bothun GLSB Das Mapelli Weij Loebman SungWon Kwak Scd Sd
% LocalWords:  Schiminovich Iodice Belokurov Fairchild RMS pPXF Cappellari th
% LocalWords:  GANDALF Sarzi CosmoloPy Bernardi GLSBs DeBlok Verheijen Yoachim
% LocalWords:  SEDs Brookhaven Instituto Astrofisica Canarias Notre JINA DM
% LocalWords:  IFU UGC HSB SAURON ellipticals Spitzer IC Galaz Myr
% LocalWords:  kpc HIPASS CCD metallicity MANGA SDSS arcsec lensing IRAF
% LocalWords:  Spectrograph spectrograph IDL longslit photometric FWHM SSP mis
% LocalWords:  dispersions Balmer decrement fiducia redshift Chabrier
% LocalWords:  Faber redshifts NGC kinemetry photometry anisotropy Gyr SFH
% LocalWords:  metallicities colliders rotators et al MNRAS de datacube Ppxf
% LocalWords:  dashed CDM Kasparova LSB Dutton Raskutti Sa Morganti
%%  LocalWords:  Emonts Weijmans Ibata McConnachie Schweizer Magain
%%  LocalWords:  Oosterloo baryonic Buote GALEX Salim Kelson outlier
%%  LocalWords:  Bruzual Ahn Willett SQL Chilingarian Saulder Swaters
%%  LocalWords:  Krajn GN Kuntschner situ indices Emsell Capp CasJobs
%%  LocalWords:  Schmitz circ accrete jabref CMD collisional GLSB's
%%  LocalWords:  impactor Conroy brightnesses Malmquist dex Hagen
%%  LocalWords:  Dmitry isothermal
